---
title: "District Statistics 2011 analysis"
author: "Tristan Kindig"
date: "January 19, 2019"
output:
  html_notebook:
    toc: yes
---

```{r include = FALSE}
# import libraries
library(ggrepel)
library(kableExtra)
library(purrr)
library(sf)
library(tidyverse)

# load CSV data
dredging_plan_df <- read_csv("./data/csvs/dredging_plan.csv")
district_statistics_df <- read_csv("./data/csvs/district_statistics.csv")

# load shapefile data
bgd_nation_sf <- st_read("./data/shapefiles/bgd_nation/bgd_nation.shp")
bgd_divisions_sf <- st_read("./data/shapefiles/bgd_divisions/bgd_divisions.shp")
bgd_districts_sf <- st_read("./data/shapefiles/bgd_districts/bgd_districts.shp")
bgd_hydrology_sf <- st_read("./data/shapefiles/bgd_hydrology/bgd_hydrology.shp")
bgd_iwt_routes_sf <- st_read("./data/shapefiles/bgd_iwt_routes/bgd_iwt_routes.shp")
bgd_ports_sf <- st_read("./data/shapefiles/bgd_ports/bgd_ports.shp")

# add population density, midpoint temperature, riverine area percentage, agricultural output, and waterways/sq_km to district_statistics_df
district_statistics_df <- district_statistics_df %>%
  mutate(
    density_pop_per_sq_km = round(population / total_area_sq_km),
    midpoint_temperature = round(
      (max_temp_c + min_temp_c) / 2,
      1
    ),
    riverine_area_percentage = round(riverine_area_sq_km / total_area_sq_km * 100),
    agricultural_product_metric_ton = (rice_metric_ton + wheat_metric_ton + sugarcane_metric_ton + potato_metric_ton),
    monsoon_waterway_km_per_sq_km = round(waterway_monsoon_season_km / total_area_sq_km)
  ) %>%
  arrange(district)

# remove unnecessary attributes from bgd_districts_sf
undesired_columns <- c("ADM2_PCODE", "ADM2_REF", "ADM2ALT1EN", "ADM2ALT2EN", "ADM1_EN", "ADM1_PCODE", "ADM0_EN", "ADM0_PCODE", "date", "validOn", "ValidTo")
bgd_districts_sf <- bgd_districts_sf %>%
  modify_at(undesired_columns, ~NULL)

# convert factor attributes to character attributes
bgd_divisions_sf$ADM1_EN <- as.character(bgd_divisions_sf$ADM1_EN)
bgd_districts_sf$ADM2_EN <- as.character(bgd_districts_sf$ADM2_EN)
bgd_ports_sf$River_Port <- as.character(bgd_ports_sf$River_Port)

# create division_statistics_df
div_stats_part_1 <- district_statistics_df %>%
  group_by(division) %>%
  summarize_at(
    c("total_area_sq_km", "riverine_area_sq_km", "population", "rainfall_meters", "rice_metric_ton", "wheat_metric_ton", "sugarcane_metric_ton", "potato_metric_ton", "road_km", "railway_km", "waterway_monsoon_season_km", "waterway_year_round_km", "railway_station", "steamer_station", "bus_station"),
    sum,
    na.rm = TRUE
  )
div_stats_part_2 <- district_statistics_df %>%
  group_by(division) %>%
  summarize_at(
    c("max_temp_c", "min_temp_c", "midpoint_temperature"),
    median,
    na.rm = TRUE
  )
division_statistics_df <- left_join(div_stats_part_1, div_stats_part_2, by = "division")

# add population density, midpoint temperature, riverine area percentage, total agricultural output, and waterways/sq_km to division_statistics_df
division_statistics_df <- division_statistics_df %>%
  mutate(
    density_pop_per_sq_km = round(population / total_area_sq_km),
    riverine_area_percentage = round(riverine_area_sq_km / total_area_sq_km * 100),
    agricultural_product_metric_ton = (rice_metric_ton + wheat_metric_ton + sugarcane_metric_ton + potato_metric_ton),
    monsoon_waterway_km_per_sq_km = round(waterway_monsoon_season_km / total_area_sq_km, 1)
  ) %>%
  arrange(division)

# join df to sf
bgd_divisions_sf <- left_join(bgd_divisions_sf, division_statistics_df, by=c("ADM1_EN" = "division"))
bgd_districts_sf <- left_join(bgd_districts_sf, district_statistics_df, by=c("ADM2_EN" = "district"))

# rename Pangaon Port to Pangaon
bgd_ports_sf <- bgd_ports_sf %>% 
    mutate(River_Port = replace(River_Port, River_Port == "Pangaon Port", "Pangaon"))
```



## Population

### Population by division (column chart)
```{r echo = FALSE}
ggplot(division_statistics_df) +
  geom_col(
    aes(
      reorder(division, population),
      population/1000000
    ),
    fill = "red"
  ) +
  geom_text(
    aes(
      x = division, 
      y = population/1000000 + 1,
      label = round(population/1000000, 1)
    )
  ) +
  labs(
    x = "Division",
    y = "Population (mil)"
  )
```

### Population density by division (chloropleth map)
```{r echo = FALSE}
ggplot(bgd_divisions_sf) +
  geom_sf(
    aes(fill = density_pop_per_sq_km)
  ) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    fill = "Population density",
    x = "Longitude",
    y = "Latitude"
  )
```


### Population density by district (chloropleth map)
```{r echo = FALSE}
bgd_districts_sf_minus_narayanganj <- bgd_districts_sf %>%
  filter(
    density_pop_per_sq_km < 4000
  )

ggplot(bgd_districts_sf_minus_narayanganj) +
  geom_sf(
    aes(fill = density_pop_per_sq_km)
  ) +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = "Population density",
    subtitle = "Narayanganj (4300) removed"
  )
```

### Population density by district (percentage histogram)
```{r echo = FALSE}
ggplot(district_statistics_df, aes(x = population/total_area_sq_km)) +  
  geom_histogram(
    aes(
      y = (..count..)/sum(..count..)
    ),
    breaks = seq(0, 2000, 200),
    fill = "red"
  ) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(0, 2000, 200)) +
  labs(
    subtitle = "n = 63; 1 outlier (Narayanganj, 4300) removed",
    x = "Population density",
    y = "Percentage"
  )
```



## Meterology

### Temp by division
```{r echo = FALSE}
ggplot(bgd_divisions_sf) +
  geom_sf(
    aes(fill = midpoint_temperature)
  ) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = median(division_statistics_df$midpoint_temperature)
  ) +
  labs(
    fill = "Temperature",
    x = "Longitude",
    y = "Latitude"
  )
```

### Temp by district
```{r echo = FALSE}
ggplot(bgd_districts_sf) +
  geom_sf(
    aes(fill = midpoint_temperature)
  ) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = median(district_statistics_df$midpoint_temperature)
  ) +
  labs(
    fill = "Temperature",
    x = "Longitude",
    y = "Latitude"
  )
```

### Annual rainfall, by district
```{r echo = FALSE}
ggplot(bgd_districts_sf) +
  geom_sf(
    aes(fill = rainfall_meters / total_area_sq_km)
  ) +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = expression(
      paste(
        "Annual rainfall (",
        frac(
          "m",
          paste("km"^"2")
        ),
        ")"
      )
    )
  )
```



## Hydrology

### Hydrological map of Bangladesh
```{r echo = FALSE, results = "hide", fig.keep = "all"}
ggplot() +
geom_sf(
  data = bgd_nation_sf,
  color = "light green",
  fill = "light green"
) +
geom_sf(
  data = bgd_hydrology_sf,
  color = "dark blue",
  fill = "light blue"
) +
labs(
  x = "Longitude",
  y = "Latitude"
)
```

### Riverine area percentage by district
```{r echo = FALSE}
ggplot(bgd_districts_sf) +
  geom_sf(
    aes(fill = riverine_area_percentage)
  ) +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = "Riverine area (percentage)"
  )
```

### Monsoon season waterways
```{r echo = FALSE}
ggplot(bgd_districts_sf) +
  geom_sf(
    aes(fill = (waterway_monsoon_season_km/total_area_sq_km))
  ) +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = expression(
      paste(
        "Monsoon season waterway length (",
        frac(
          "km",
          paste("km"^"2")
        ),
        ")"
      )
    )
  )
```



## Agriculture

### Agricultural production by division (chloropleth map)
```{r echo = FALSE}
ggplot(bgd_divisions_sf) +
  geom_sf(
    aes(fill = agricultural_product_metric_ton/total_area_sq_km)
  ) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = expression(
      paste(
        "Agricultural production (",
        frac(
          "metric tons",
          paste("km"^"2")
        ),
        ")"
      )
    )
  )
```

### Agricultural production by district (chloropleth map)
```{r echo = FALSE}
ggplot(bgd_districts_sf) +
  geom_sf(
    aes(fill = agricultural_product_metric_ton/total_area_sq_km)
  ) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(
    x = "Longitude",
    y = "Latitude",
    fill = expression(
      paste(
        "Agricultural production (",
        frac(
          "metric tons",
          paste("km"^"2")
        ),
        ")"
      )
    )
  )
```

### Agricultural production by district (percentage histogram)
```{r echo = FALSE}
ggplot(district_statistics_df, aes(x = agricultural_product_metric_ton/total_area_sq_km)) +  
  geom_histogram(
    aes(
      y = (..count..)/sum(..count..)
    ),
    breaks = seq(0, 1750, 175),
    fill = "green",
    na.rm = TRUE
  ) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(0, 1750, 175)) +
  labs(
    x = "Agr product mt / sq_km",
    y = "Percentage"
  )
```



## Industry



## Transportation

### Rivers and ports (chloropleth map, zoomed out)
```{r echo = FALSE, results = "hide", fig.keep = "all"}
ggplot() +
geom_sf(
  data = bgd_nation_sf,
  color = "light green",
  fill = "light green"
) +
geom_sf(
  data = bgd_iwt_routes_sf,
  color = "blue",
) +
geom_sf(
  data = bgd_ports_sf,
  color = "red",
  size = 0.7
) +
labs(
  x = "Longitude",
  y = "Latitude"
)
```

### Rivers and ports (chloropleth map, zoomed in)
```{r echo = FALSE, results = "hide", fig.keep = "all"}
ggplot() +
geom_sf(
  data = bgd_nation_sf,
  color = "light green",
  fill = "light green"
) +
geom_sf(
  data = bgd_iwt_routes_sf,
  color = "blue",
) +
geom_sf(
  data = bgd_ports_sf,
  color = "red",
  size = 1.5
) +
geom_text_repel(
  data = bgd_ports_sf,
  aes(
    x = bgd_ports_sf$Longitude,
    y = bgd_ports_sf$Latitude,
    label = bgd_ports_sf$River_Port,
    geometry = geometry
  ),
  stat = "sf_coordinates",
  min.segment.length = 0
) +
labs(
  x = "Longitude",
  y = "Latitude"
) +
coord_sf(
  xlim = c(89, 92),
  ylim = c(21.5, 24.25)
)
```



## Data sources

**dredging information**: https://vanderbilt.app.box.com/file/382909402931

**hydrological map shapefile**: https://geonode.wfp.org/layers/geonode:bgd_hyd_major_rivers_py

**port shapefile**: MacKenzie

**river network shapefile**: Amir

**administrative boundary shapefiles**: https://data.humdata.org/dataset/administrative-boundaries-of-bangladesh-as-of-2015

**riverine area, monsoon waterway length, and annual rainfall**: http://203.112.218.65:8008/PageWebMenuContent.aspx?MenuKey=246