---
title: "Dredging plan analysis"
author: "Tristan Kindig"
date: "January 24, 2019"
output:
  html_notebook:
    toc: yes
---

```{r include = FALSE}
# import libraries
library(ggrepel)
library(kableExtra)
library(purrr)
library(sf)
library(tidyverse)

# load CSV data
dredging_plan_df <- read_csv("./data/csvs/dredging_plan.csv")
district_statistics_df <- read_csv("./data/csvs/district_statistics.csv")

# load shapefile data
bgd_nation_sf <- st_read("./data/shapefiles/bgd_nation/bgd_nation.shp")
bgd_divisions_sf <- st_read("./data/shapefiles/bgd_divisions/bgd_divisions.shp")
bgd_districts_sf <- st_read("./data/shapefiles/bgd_districts/bgd_districts.shp")
bgd_hydrology_sf <- st_read("./data/shapefiles/bgd_hydrology/bgd_hydrology.shp")
bgd_iwt_routes_sf <- st_read("./data/shapefiles/bgd_iwt_routes/bgd_iwt_routes.shp")
bgd_ports_sf <- st_read("./data/shapefiles/bgd_ports/bgd_ports.shp")

# add population density, riverine area percentage, agricultural output, and waterways/sq_km to district_statistics_df
district_statistics_df <- district_statistics_df %>%
  mutate(
    density_pop_per_sq_km = round(population / total_area_sq_km),
    midpoint_temperature = round(
      (max_temp_c + min_temp_c) / 2,
      1
    ),
    riverine_area_percentage = round(riverine_area_sq_km / total_area_sq_km * 100),
    agricultural_product_metric_ton = round(
      (rice_metric_ton + wheat_metric_ton + sugarcane_metric_ton + potato_metric_ton),
      1
    ),
    monsoon_waterway_km_per_sq_km = round(waterway_monsoon_season_km / total_area_sq_km)
  ) %>%
  arrange(district)

# remove unnecessary attributes from bgd_districts_sf
undesired_columns <- c("ADM2_PCODE", "ADM2_REF", "ADM2ALT1EN", "ADM2ALT2EN", "ADM1_EN", "ADM1_PCODE", "ADM0_EN", "ADM0_PCODE", "date", "validOn", "ValidTo")
bgd_districts_sf <- bgd_districts_sf %>%
  modify_at(undesired_columns, ~NULL)

# convert factor attributes to character attributes
bgd_divisions_sf$ADM1_EN <- as.character(bgd_divisions_sf$ADM1_EN)
bgd_districts_sf$ADM2_EN <- as.character(bgd_districts_sf$ADM2_EN)
bgd_ports_sf$River_Port <- as.character(bgd_ports_sf$River_Port)

# add population density, midpoint temperature, riverine area percentage, agricultural output, and waterways/sq_km to district_statistics_df
district_statistics_df <- district_statistics_df %>%
  mutate(
    density_pop_per_sq_km = round(population / total_area_sq_km),
    midpoint_temperature = round(
      (max_temp_c + min_temp_c) / 2,
      1
    ),
    riverine_area_percentage = round(riverine_area_sq_km / total_area_sq_km * 100),
    agricultural_product_metric_ton = round(
      (rice_metric_ton + wheat_metric_ton + sugarcane_metric_ton + potato_metric_ton),
      2
    ),
    monsoon_waterway_km_per_sq_km = round(waterway_monsoon_season_km / total_area_sq_km)
  ) %>%
  arrange(district)

# create division_statistics_df
div_stats_part_1 <- district_statistics_df %>%
  group_by(division) %>%
  summarize_at(
    c("total_area_sq_km", "riverine_area_sq_km", "population", "rainfall_meters", "rice_metric_ton", "wheat_metric_ton", "sugarcane_metric_ton", "potato_metric_ton", "road_km", "railway_km", "waterway_monsoon_season_km", "waterway_year_round_km", "railway_station", "steamer_station", "bus_station"),
    sum,
    na.rm = TRUE
  )
div_stats_part_2 <- district_statistics_df %>%
  group_by(division) %>%
  summarize_at(
    c("max_temp_c", "min_temp_c", "midpoint_temperature"),
    median,
    na.rm = TRUE
  )
division_statistics_df <- left_join(div_stats_part_1, div_stats_part_2, by = "division")

# add population density, riverine area percentage, total agricultural output, and waterways/sq_km to division_statistics_df
division_statistics_df <- division_statistics_df %>%
  mutate(
    density_pop_per_sq_km = round(population / total_area_sq_km),
    riverine_area_percentage = round(riverine_area_sq_km / total_area_sq_km * 100),
    agricultural_product_metric_ton = round(
      (rice_metric_ton + wheat_metric_ton + sugarcane_metric_ton + potato_metric_ton),
      2
    ),
    monsoon_waterway_km_per_sq_km = round(waterway_monsoon_season_km / total_area_sq_km)
  ) %>%
  arrange(division)

# join df to sf
bgd_divisions_sf <- left_join(bgd_divisions_sf, division_statistics_df, by=c("ADM1_EN" = "division"))
bgd_districts_sf <- left_join(bgd_districts_sf, district_statistics_df, by=c("ADM2_EN" = "district"))

# rename Pangaon Port to Pangaon
bgd_ports_sf <- bgd_ports_sf %>% 
    mutate(River_Port = replace(River_Port, River_Port == "Pangaon Port", "Pangaon"))
```

### Density plot of Phase 1 dredging project costs (outliers removed)
```{r echo = FALSE}
projects_under_750 <- dredging_plan_df %>%
  filter(cost_mil_usd < 750)

ggplot(projects_under_750) +
  geom_density(
    aes(x = cost_mil_usd),
    fill = "red"
  ) +
  geom_vline(
    xintercept = median(projects_under_750$cost_mil_usd),
    color = "black"
  ) +
  annotate(
    "text",
    x = median(projects_under_750$cost_mil_usd),
    y = 0.003,
    angle = 90,
    vjust = 1.5,
    hjust = 0.9,
    color = "black",
    size = 4,
    label = paste("median = ", round(median(projects_under_750$cost_mil_usd)))
  ) +
  labs(
    subtitle = "n = 219; 4 outliers removed",
    x = "Cost (million USD)",
    y = "Density"
  ) +
  xlim(0, 750)
```

### Density plot of Phase 1 dredging project costs by agency (outliers removed)
```{r echo = FALSE}
ggplot(projects_under_750) +
  geom_density(
    aes(x = cost_mil_usd, fill = agency, color = agency),
    alpha = 0.3
  ) +
  labs(
    x = "Cost (million USD)",
    y = "Density"
  )
```

### Histogram of Phase 1 dredging project costs, stacked by agency
```{r echo = FALSE}
ggplot(
  projects_under_750,
  aes(x = cost_mil_usd)
) + 
stat_bin(
  aes(fill = agency),
  breaks = seq(0, 750, 75)
) +  
stat_bin(
  breaks = seq(0, 750, 75),
  geom="text",
  aes(label=..count..),
  vjust=-0.5
) +
scale_x_continuous(breaks = seq(0, 750, 75)) +
ylim(c(0, 150)) +
labs(
  subtitle = "10 bins, each 75 wide; 4 outliers removed; n = 219",
  x = "Cost (million USD)",
  y = "Count",
  fill = "Agency"
)
```

### Histogram of Phase 1 dredging project costs, stacked by activity
```{r echo = FALSE}
ggplot(
  projects_under_750,
  aes(x = cost_mil_usd)
) + 
stat_bin(
  aes(fill = activity),
  breaks = seq(0, 750, 75)
) +  
stat_bin(
  breaks = seq(0, 750, 75),
  geom="text",
  aes(label=..count..),
  vjust=-0.5
) +
scale_x_continuous(breaks = seq(0, 750, 75)) +
ylim(c(0, 150)) +
labs(
  subtitle = "10 bins, each 75 wide; 4 outliers removed; n = 219",
  x = "Cost (million USD)",
  y = "Count",
  fill = "Activity"
)
```

### Overall Phase 1 dredging project statistics
```{r results = "asis", echo = FALSE}
activity_stats <- dredging_plan_df %>%
  summarize(
    projects = n(),
    total_cost_mil_usd = round(sum(cost_mil_usd)),
    median_cost_mil_usd = round(median(cost_mil_usd))
  )

kable(activity_stats) %>%
  kable_styling(full_width = FALSE)
```

### Phase 1 dredging project outliers
```{r results = "asis", echo = FALSE}
projects_over_750 <- dredging_plan_df %>%
  filter(750 <= cost_mil_usd) %>%
  arrange(desc(cost_mil_usd)) %>%
  select(location, cost_mil_usd, activity, zone, agency)

projects_over_750$cost_mil_usd <- round(projects_over_750$cost_mil_usd)

kable(projects_over_750) %>%
  kable_styling(full_width = FALSE)
```

### Phase 1 dredging project statistics, by agency
```{r results = "asis", echo = FALSE}
agency_stats <- dredging_plan_df %>%
  group_by(agency) %>%
  summarize(
    projects = n(),
    total_cost_mil_usd = round(sum(cost_mil_usd)),
    median_cost_mil_usd = round(median(cost_mil_usd))
  ) %>%
  arrange(desc(projects))

kable(agency_stats) %>%
  kable_styling(full_width = FALSE)
```

### Phase 1 dredging project statistics, by activity
```{r results = "asis", echo = FALSE}
activity_stats <- dredging_plan_df %>%
  group_by(activity) %>%
  summarize(
    projects = n(),
    total_cost_mil_usd = round(sum(cost_mil_usd)),
    median_cost_mil_usd = round(median(cost_mil_usd))
  )

kable(activity_stats) %>%
  kable_styling(full_width = FALSE)
```